---
title: "HW1_exer2"
author: "Group9"
date: "9/4/2020"
output: html_document
---

# Exercise 2: Financial Analysis

First we need to make sure we have all we need:

```{r trueinstall, results=FALSE, message=FALSE, warning=FALSE}

if (!require("quantmod")) {
    install.packages("quantmod")
    library(quantmod)
}

```

The data set from four companies was obtained from yahoo, from January 1-2019 to September 6-2020. This was done utilizing the package `quantmod`:

```{r install, results=FALSE, message=FALSE, warning=FALSE}

start <- as.Date("2019-01-01")
end <- as.Date(Sys.Date()) 

getSymbols(c("VALE", "CSCO", "INTC", "WFC"), src="yahoo", from=start, to=end)

```

After that, all the data sets should be fetched and can be seen in the environment tab. We can take a look at them here using the command `head(dataset)`. For the VALE dataset, for example:

```{r viewdata, include=TRUE, eval=TRUE, echo=FALSE}

head(VALE)

```

An xts object `stocks` was created to unify all closing values from all the datasets, and then used to plot these values:

```{r stocks_object, include=TRUE, eval=TRUE}

stocks <- as.xts(data.frame(VALE=VALE[, "VALE.Close"],
                            CSCO=CSCO[, "CSCO.Close"],
                            INTC=INTC[, "INTC.Close"],
                            WFC=WFC[, "WFC.Close"]
                            ))
```

```{r plot_, echo=TRUE, results=FALSE, fig.show='hide', message=FALSE, warning=FALSE}

plot(stocks,
     screens=1,
     lty=1:4,
     xlab="Date",
     ylab="Price",
     font.axis=1)
```

```{r plot, fig.align="center", fig.width=8, fig.cap="Figure 1. Chart of the closing price of the four stocks over the period of January 01, 2019 to September 06, 2020.", include=TRUE, eval=TRUE}

addLegend(legend.loc = "topleft",
          legend.names = c("Vale S.A.", "Cisco", "Intel", "Wells Fargo"),
          lty= 1:4,
          col = NULL,
          ncol = 1,
          cex = 0.8,
          on = 1)
```
<!--
```{r candlechart, fig.align="center", fig.width=8, fig.cap="candlestick chart of the four stocks over the last four months", include=TRUE, eval=TRUE}

FinancialAnalysis <- as.xts(data.frame(VALE=VALE['2020-05-01/2020-09-06'],
                                       CSCO=CSCO['2020-05-01/2020-09-06'],
                                       INTC=INTC['2020-05-01/2020-09-06'],
                                       WFC=WFC['2020-05-01/2020-09-06']
                                       ))
candleChart(FinancialAnalysis,
            up.col="gray",
            dn.col="blue",
            theme="white")


```
-->


### 4
### 5
### 6
### 7 
