---
title: "HW1_exer2"
author: "Group9"
date: "9/4/2020"
output: html_document
---

# Exercise 2: Financial Analysis
\

First we need to a package:

```{r trueinstall, results=FALSE, message=FALSE, warning=FALSE}

if (!require("quantmod")) {
    install.packages("quantmod")
    library(quantmod)
}

```
\

Then, we need to fetch the data. This data set from four companies will obtained from yahoo, from January 1-2019 to today's date. This is done utilizing the package `quantmod`:

```{r dataset, results=FALSE, message=FALSE, warning=FALSE}

start <- as.Date("2019-01-01")
end <- as.Date(Sys.Date()) 

getSymbols(c("VALE", "CSCO", "INTC", "WFC"), src="yahoo", from=start, to=end)

```
\

After that, all the data sets should be fetched and can be seen in the environment tab. We can take a look at them here using the command `head(dataset)`. For the VALE dataset, for example:

```{r viewdata, include=TRUE, eval=TRUE, echo=FALSE}

head(VALE)

```
\

An xts object `stocks` will created to unify all closing values from all the datasets, and then used to plot these values:

```{r stocks_object, include=TRUE, eval=TRUE}

stocks <- as.xts(data.frame(VALE=VALE[, "VALE.Close"],
                            CSCO=CSCO[, "CSCO.Close"],
                            INTC=INTC[, "INTC.Close"],
                            WFC=WFC[, "WFC.Close"]
                            ))
```

```{r plot_, echo=TRUE, results=FALSE, fig.show='hide', message=FALSE, warning=FALSE}

plot(stocks,
     screens=1,
     lty=1:4,
     xlab="Date",
     ylab="Price",
     font.axis=1)
```

```{r plot, fig.align="center", fig.width=8, fig.cap="Figure 1. Chart of the closing price of the four stocks over the period of January 01, 2019 to September 06, 2020.", include=TRUE, eval=TRUE}

addLegend(legend.loc = "topleft",
          legend.names = c("Vale S.A.", "Cisco", "Intel", "Wells Fargo"),
          lty= 1:4,
          col = NULL,
          ncol = 1,
          cex = 0.8,
          on = 1)
```
\

We can also plot a Candle Chart of the four stocks over the last four months:

```{r candlechart, fig.align="center", fig.width=8, fig.cap="Figure 2. Candlestick chart of the four stocks over the past four months.", include=TRUE, eval=TRUE}

# Redefining the data range and the data object
startC <- as.Date("2020-05-01")
FinancialAnalysis <- as.xts(data.frame(VALE=VALE['2020-05-01/2020-09-06'],
                                       CSCO=CSCO['2020-05-01/2020-09-06'],
                                       INTC=INTC['2020-05-01/2020-09-06'],
                                       WFC=WFC['2020-05-01/2020-09-06']
))

# Calling the plot and making it nicer
candleChart(FinancialAnalysis,
            up.col="gray",
            dn.col="blue",
            theme="white")
```
\

Each day is represented by a "candle" which appears to be similar to a box plot, but the anatomy of a candle is slightly different.

**Wicks:** First, the lines or "wicks" sticking out of the top and bottom represent the high and low price of the stock. The top wick represents the highest price reached for the day while the bottom wick represents the lowest price reached.

**Color:** The color of the box represents if the price increased or decreased from the beginning to the end of the day. <font color="gray">Gray</font> boxes represent an overall increase for the day while <font color="blue">blue</font> boxes represent a decrease.

**Boxes:** Depending on the color, the top and bottom of the box (excluding the wick) changes meaning. If the box is <font color="gray">gray</font> representing an increase, then the bottom of the box is the opening price of the stock and the top of the box is the closing price. If the box is <font color="blue">blue</font> representing an overall decrease, then the top of the box is the closing price while the bottom is the opening price.

\

To compute the mean and variance of all four time series using the closing price of the stock, we use:

```{r mean_var, include=TRUE, eval=TRUE, results='markup', warning=FALSE}

#applying the Mean and Variance functions with a vector output into a data.frame object
stocks_mean <- data.frame(sapply(stocks, mean)) 
stocks_var <- data.frame(sapply(stocks, var))

# put the results together
stocks_stats <- cbind(stocks_mean, stocks_var) 

#making it prettier/intelligible
colnames(stocks_stats) <- c("Mean", "Variance")
rownames(stocks_stats) <- c("Vale S.A.", "Cisco", "Intel", "Wells Fargo")

```     
\

<!--
dimnames(stocks_stats)[names(stocks_stats) == "VALE.Close"] <- "Vale S.A."
dimnames(stocks_stats)[names(stocks_stats) == "CSCO.Close"] <- "Cisco"
dimnames(stocks_stats)[names(stocks_stats) == "INTC.Close"] <- "Intel"
dimnames(stocks_stats)[dimnames(stocks_stats) == "WFC.Close"] <- "Wells Fargo"
-->

Voilà
```{r knitr, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

if (!require("knitr")) {
     install.packages("knitr")
     library(knitr)
}

kable(stocks_stats)
```
From these means we can see Vale S.A. has a much lower average price over the past four months when compared to the other three, and Intel Corporation has the highest average stock price.

The mean stock price doesn't tell the whole story, however. It is a good idea to calculate the variance which indicates how spread out the data are from the mean. A higher stock price variance indicates a higher volatility in the past four months. The variances of each stock are calculated below. From them we can see that although Vale S.A. had the lowest average stock price over the past four months, it also has the lowest variance, indicating that it stayed the most consistent. Wells Fargo, on the other hand, had the largest variance by a large margin, indicating that its stock price was the most volatile out of the four.

\

<!--
WHY THIS DOES NOT WORK?
stock_stats_kable <- knitr::kable(head(stocks_stats), "html")
kable_styling(stock_stats_kable, full_width=FALSE, position="center")
-->

Next, we want to compare these 4 stocks by their yearly Sharpe ratio \(S\) based on the last 10 years, with an yearly risk-free rate of return \(Rf\) is of \(3.00\:\%\): First we need to fetch the data for the last 10 years.

```{r dataset_year, results=FALSE, message=FALSE, warning=FALSE}

start <- as.Date("2010-09-01") #because I'm dumb
end <- as.Date(Sys.Date()) 

getSymbols(c("VALE", "CSCO", "INTC", "WFC"),
           src="yahoo",
           from=start,
           to=end)

```
Now.. Math!

The daily Sharpe Ratio is related to the yearly Sharpe Ratio by \(\tilde{S} = \frac{S}{\sqrt{252}}\), for which \(\tilde{S}\) is defined as
\(\tilde{S} = \frac{\mathbb{E}[R_t - \tilde{R}_f]}{\sigma_{\mu}}\), where:
\(\tilde{S}\) is the daily Sharpe ratio;
\(\mathbb{E}[\cdot]\) stands for the expected value (estimated by the mean);
\(R_t\) is the asset return defined as \(R_t = \frac{X_t - X_{t-1}}{X_{t-1}}\) where \(X_t\) is the price of an asset at time \(t\);
\(\tilde{R}_f\) is the daily risk-free rate of return defined as \(\tilde{R}_f = \frac{R_f}{252}\); and
\(\sigma_{\mu}\) is the standard deviation of the asset return.

```{r sharpe, eval=TRUE, message=FALSE, warning=FALSE}

# assign common variables
Rf <- 3/100
Rf_ <- Rf/252

# Assign specific variables (see text full equations)
AssetReturnVALE <- dailyReturn(VALE)
stdVALE <- sd(AssetReturnVALE)
meanVALE <- mean(AssetReturnVALE)

AssetReturnCSCO <- dailyReturn(CSCO)
stdCSCO <- sd(AssetReturnCSCO)
meanCSCO <- mean(AssetReturnCSCO)

AssetReturnINTC <- dailyReturn(INTC)
stdINTC <- sd(AssetReturnINTC)
meanINTC <- mean(AssetReturnINTC)

AssetReturnWFC <- dailyReturn(WFC)
stdWFC <- sd(AssetReturnWFC)
meanWFC <- mean(AssetReturnWFC)

# Calculating Yearly Sharpe Ratio based on Daily Sharpe Ratio:

# Vale S.A.
S_VALE <- (meanVALE - Rf_)/stdVALE
SharpeVALE <- S_VALE * sqrt(252)

# Cisco
S_CSCO <- (meanCSCO - Rf_)/stdCSCO
SharpeCSCO <- S_CSCO * sqrt(252)

# Intel
S_INTC <- (meanINTC - Rf_)/stdINTC
SharpeINTC <- S_INTC * sqrt(252)

#Wells Fargo
S_WFC <- (meanWFC - Rf_)/stdWFC
SharpeWFC <- S_WFC * sqrt(252)
```

```{r sharpeshow, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}

if (!require("knitr")) {
     install.packages("knitr")
     library(knitr)
}

CalculatedSharpeRatio <- c(SharpeVALE, SharpeCSCO, SharpeINTC, SharpeWFC)
Company <- c("Vale S.A.", "Cisco", "Intel", "Wells Fargo")

SharpeResults <- data.frame(Company, CalculatedSharpeRatio)

kable(SharpeResults)

```
\

<!--
I think this is right but not sure? Equally weighted just means adding across all rows then dividing by 4 to make a new column

Then mean & variance of that column
-->

Now we consider consider an equally weighted portfolio of all of these stocks combined. We simply add each of the four stocks together per day, then take the average to calculate the equally weighted new stock. Below are the four stocks listed again with the "Combined" stock at the bottom for comparison:
     
```{r dataset_EW, echo=FALSE, eval=TRUE, results=FALSE, message=FALSE, warning=FALSE}

start <- as.Date("2010-09-01") #because I'm dumb
end <- as.Date(Sys.Date()) 

getSymbols(c("VALE", "CSCO", "INTC", "WFC"),
           src="yahoo",
           from=start,
           to=end)

```

```{r EqualWeight, eval=TRUE, message=FALSE, warning=FALSE}

n <- ncol(stocks)

weights <- rep(1/n,n)

weights <- stocks / sum(data.frame(stocks))


```
```{r combined, include=TRUE, eval=TRUE, message=FALSE, warning=FALSE}

stocks$Totals <- rowSums(stocks[,c("VALE.Close", "CSCO.Close", "INTC.Close", "WFC.Close")])/4

#applying the Mean and Variance functions with a vector output into a data.frame object
stocks_mean <- data.frame(sapply(stocks, mean)) 
stocks_var <- data.frame(sapply(stocks, var))

# put the results together
stocks_stats <- cbind(stocks_mean, stocks_var) 

#making it prettier/intelligible
colnames(stocks_stats) <- c("Mean", "Variance")
rownames(stocks_stats) <- c("Vale S.A.", "Cisco", "Intel", "Wells Fargo", "Combined")

```

```{r knitr2, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

if (!require("knitr")) {
     install.packages("knitr")
     library(knitr)
}

kable(stocks_stats)
```
Since they were equally weighted, the mean of the Combined stock is just an average of the averages. When looking at variance, it is the second lowest, only having a higher variance than Vale S.A.

<!--
EWreturn <- as.matrix(rep(1/length(stocks),length(stocks)))
resEqual = as.matrix(returns) %*% t(EWreturn)
colnames(resEqual) = “EqualWeight Portfolio”
-->