---
title: "HW1_exer2"
author: "Group 9: Ana Gabriela Itokazu, Cary Burdick, Harsha Galkanda-Arachchige"
date: "Final Version: September 18, 2020."
output: html_document
bibliography: refs2.bib
---
<a href="https://github.com/AU-R-Programming/HW1-9/blob/master/HW1_exer2.Rmd">source</a>


# Exercise 2: Financial Analysis 
\

<div style="text-align: justify">

 
First things first,

```{r trueinstall, results=FALSE, message=FALSE, warning=FALSE}

if (!require("quantmod")) {
    install.packages("quantmod")
    library(quantmod)
}

```
\

 
Then, we need to fetch the data. This data set from four companies will obtained from Yahoo, from January 1-2019 to today's date. This is done utilizing the package `quantmod`:

```{r dataset, results=FALSE, message=FALSE, warning=FALSE}

start <- as.Date("2019-01-01")
end <- as.Date(Sys.Date()) 

getSymbols(c("VALE", "CSCO", "INTC", "WFC"),
           src="yahoo",
           from=start,
           to=end)

```
\

 
After that, the data sets should be fetched and can be seen in the environment tab. The first and last lines for the `VALE` dataset, for example:

```{r knitr, include=TRUE, eval=TRUE, echo=FALSE, message=FALSE}

if (!require("knitr")) {
     install.packages("knitr")
     library(knitr)
}

VALEview <- rbind(data.frame(first(VALE,n=1,keep=FALSE)),
                  data.frame(last(VALE,n=1,keep=FALSE)))

colnames(VALEview) <- c("Open", "High", "Low", "Close", "Volume", "Adjusted")

kable(VALEview)

```
\

 
We can create an `xts` object called `stocks` to unify all closing values from all the datasets (this `eXtensible Time Series` object will be used in several steps of this exercise). Then we can use the `plot()` function to take a look on it, and add a good-looking legend using the `addLegend()` function:

```{r stocks_object, include=TRUE, eval=TRUE}

stocks <- as.xts(data.frame(VALE=VALE[, "VALE.Close"],
                            CSCO=CSCO[, "CSCO.Close"],
                            INTC=INTC[, "INTC.Close"],
                            WFC=WFC[, "WFC.Close"]
                            ))
```

```{r plot_, echo=TRUE, results=FALSE, fig.show='hide', message=FALSE, warning=FALSE}

plot(stocks,
     screens=1,
     lty=1:4,
     xlab="Date",
     ylab="Price",
     font.axis=1)
```

```{r plot, fig.align="center", fig.width=8, fig.cap="Figure 1. Chart of the closing price of the four stocks over the period of January 01, 2019 to September 06, 2020.", include=TRUE, eval=TRUE}

addLegend(legend.loc = "topleft",
          legend.names = c("Vale S.A.", "Cisco", "Intel", "Wells Fargo"),
          lty= 1:4,
          col = NULL,
          ncol = 1,
          cex = 0.8,
          on = 1)
```
\

From the graph, we can see that the closing price of Vale S.A. is much lower than that of Cisco, Intel, and Wells Fargo throughout the year. It's also very easy to see the sharp decline in the closing price of each stock in early March of 2020 indicating the start of the COVID-19 recession. Vale S.A.'s price began increasing again in May and has been steadily climbing, though it has not reached its pre-COVID price as of yet. Intel and Cisco showed signs of recovery after March with increasing prices, but both took another dip in price in July and August, respectively. Wells Fargo has shown almost no signs of recovery with its stock price sitting at about half the price of what it was before the initial hit in March.

\
 
We can also plot a Candle Chart of the four stocks over the last four months using the `candleChart()` function over a newly made `xts` object called `FinancialAnalysis`, which includes the values for all four stocks over the last 4 months:

```{r candlechart, fig.align="center", fig.width=8, fig.cap="Figure 2. Candlestick chart of the four stocks over the past four months.", include=TRUE, eval=TRUE}

# Redefining the data range and the data object

FinancialAnalysis <- as.xts(data.frame(VALE=VALE['2020-05-01/2020-09-06'],
                                       CSCO=CSCO['2020-05-01/2020-09-06'],
                                       INTC=INTC['2020-05-01/2020-09-06'],
                                       WFC=WFC['2020-05-01/2020-09-06']
))

# Calling the plot and making it nicer with some colours
candleChart(FinancialAnalysis,
            up.col="gray",
            dn.col="blue",
            theme="white")
```
\

 
In this type of graph, each day is represented by a "candle" which appears to be similar to a box plot, but the anatomy of a candle is slightly different.

 
**Wicks:** First, the lines or "wicks" sticking out of the top and bottom represent the high and low price of the stock. The top wick represents the highest price reached for the day while the bottom wick represents the lowest price reached.

 
**Color:** The color of the box represents if the price increased or decreased from the beginning to the end of the day. <font color="gray">Gray</font> boxes represent an overall increase for the day while <font color="blue">blue</font> boxes represent a decrease.

 
**Boxes:** Depending on the color, the top and bottom of the box (excluding the wick) changes meaning. If the box is <font color="gray">gray</font> representing an increase, then the bottom of the box is the opening price of the stock and the top of the box is the closing price. If the box is <font color="blue">blue</font> representing an overall decrease, then the top of the box is the closing price while the bottom is the opening price.

There are two prominent, steep decreases in mid-July and mid-August of the two highest price stocks of Intel and Cisco. According to @intel2020, for Intel specifically, July marked the announcement of the delay of its next generation 7 nano-meter computer chips which did not sit well with investors.

\
 
Coming back to that `stocks` object, we compute the mean and variance:

```{r mean_var, include=TRUE, eval=TRUE, results='markup', warning=FALSE}

#applying the Mean and Variance functions with a vector output into a data.frame object
stocks_stats <- cbind(data.frame(sapply(stocks, mean)),
                      data.frame(sapply(stocks, var))) 

#making it prettier/intelligible
colnames(stocks_stats) <- c("Mean", "Variance")
rownames(stocks_stats) <- c("Vale S.A.", "Cisco", "Intel", "Wells Fargo")
```

```{r knitr1, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

if (!require("knitr")) {
     install.packages("knitr")
     library(knitr)
}

kable(stocks_stats)
```

 
From these means we can see Vale S.A. has a much lower average price over the past four months when compared to the other three, and Intel has the highest average stock price.

 
The mean stock price doesn't tell the whole story, however. It is a good idea to calculate the variance which indicates how spread out the data are from the mean. A higher stock price variance indicates a higher volatility in the past four months. From them we can see that although Vale S.A. had the lowest average stock price over the past four months, it also has the lowest variance, indicating that it stayed the most consistent. This is evident from the candle chart as well as the candles of Vale S.A. are much smaller when compared to the other stocks indicating a smaller price spread. Wells Fargo, on the other hand, had the largest variance by a large margin, indicating that its stock price was the most volatile out of the four.

\

Next, we want to compare these 4 stocks by their yearly Sharpe ratio \(S\) based on the last 10 years, with an yearly risk-free rate of return \(Rf\) is of \(3.00\:\%\): First we need to fetch the data for the last 10 years.

```{r dataset_year, results=FALSE, message=FALSE, warning=FALSE}

start <- as.Date("2010-09-01") #because I'm dumb
end <- as.Date(Sys.Date()) 

getSymbols(c("VALE", "CSCO", "INTC", "WFC"),
           src="yahoo",
           from=start,
           to=end)

```
\

 
Now.. Math!

Consider that the daily Sharpe Ratio is related to the yearly Sharpe Ratio by 

\[\tilde{S} = \frac{S}{\sqrt{252}}\]

for which \(\tilde{S}\) is defined as

\[\tilde{S} = \frac{\mathbb{E}[R_t - \tilde{R}_f]}{\sigma_{\mu}}\]

where:

- $\tilde{S}$ is the daily Sharpe ratio;
- $\mathbb{E}[\cdot]$ stands for the expected value (estimated by the mean);
- $R_t$ is the asset return defined as $R_t = \frac{X_t - X_{t-1}}{X_{t-1}}$ where $X_t$ is the price of an asset at time $t$;
- $\tilde{R}_f$ is the daily risk-free rate of return defined as $\tilde{R}_f = \frac{R_f}{252}$; and
- $\sigma_{\mu}$ is the standard deviation of the asset return.

 
Then, we calculate (all) this for each Company, breaking down the equations and each term to be more understandable:

```{r sharpe, eval=TRUE, message=FALSE, warning=FALSE}

# assign common variables
Rf <- 3/100
Rf_ <- Rf/252

# Assign specific variables (see text full equations)
AssetReturnVALE <- dailyReturn(VALE)
stdVALE <- sd(AssetReturnVALE)
meanVALE <- mean(AssetReturnVALE)

AssetReturnCSCO <- dailyReturn(CSCO)
stdCSCO <- sd(AssetReturnCSCO)
meanCSCO <- mean(AssetReturnCSCO)

AssetReturnINTC <- dailyReturn(INTC)
stdINTC <- sd(AssetReturnINTC)
meanINTC <- mean(AssetReturnINTC)

AssetReturnWFC <- dailyReturn(WFC)
stdWFC <- sd(AssetReturnWFC)
meanWFC <- mean(AssetReturnWFC)

# Calculating Yearly Sharpe Ratio based on Daily Sharpe Ratio:

# Vale S.A.
S_VALE <- (meanVALE - Rf_)/stdVALE
SharpeVALE <- S_VALE * sqrt(252)

# Cisco
S_CSCO <- (meanCSCO - Rf_)/stdCSCO
SharpeCSCO <- S_CSCO * sqrt(252)

# Intel
S_INTC <- (meanINTC - Rf_)/stdINTC
SharpeINTC <- S_INTC * sqrt(252)

# Wells Fargo
S_WFC <- (meanWFC - Rf_)/stdWFC
SharpeWFC <- S_WFC * sqrt(252)

```
\

For which we could make a cleaner code, with the same results:

```{r sharpe2, eval=TRUE, message=FALSE, warning=FALSE}

SharpeVALE2 <- ((mean(dailyReturn(VALE))-(0.03/252))
                / sd(dailyReturn(VALE))) * sqrt(252)

SharpeCSCO2 <- ((mean(dailyReturn(CSCO))-(0.03/252))
                / sd(dailyReturn(CSCO))) * sqrt(252)

SharpeINTC2 <- ((mean(dailyReturn(INTC))-(0.03/252))
                / sd(dailyReturn(INTC))) * sqrt(252)

SharpeWFC2 <- ((mean(dailyReturn(WFC))-(0.03/252))
               / sd(dailyReturn(WFC))) * sqrt(252)

```
\

ta-da!

```{r knitr2, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

if (!require("knitr")) {
     install.packages("knitr")
     library(knitr)
}

SharpeResults <- data.frame(c("Vale S.A.", "Cisco", "Intel", "Wells Fargo"),
                            c(SharpeVALE, SharpeCSCO, SharpeINTC, SharpeWFC),
                            c(SharpeVALE2, SharpeCSCO2, SharpeINTC2, SharpeWFC2))

colnames(SharpeResults) <- c("Company", "Calculated Sharpe Ratio", "the other Calculated Sharpe Ratio")

kable(SharpeResults)

```

The Sharpe ratio is helpful in understanding the risk vs. reward of a certain stock portfolio. It takes into account the average return earned in excess of the risk-free rate (which in this case is 3%) with respect to units of volatility. The risk-free rate is a simple way of saying that investing in something with zero risk will yield a 3% return on investment (ROI). According to @sharpe2020, the higher the Sharpe ratio, the more attractive the stock is in terms of the risk adjusted reward. However, we are now taking the yearly Sharpe ratio of these four stocks from the past 10 years, so the information just received from the past four months cannot be compared directly.

It appears that Intel has the highest Sharpe ratio of the four stocks, meaning over the past 10 years it has yielded the best risk-adjusted performance. It is crucial to reiterate that this is risk-adjusted performance and not necessarily pure ROI. A stock can provide a very high return, but it may also come with very high risk and volatility, which would lower its Sharpe ratio. This is to say that Intel didn't necessarily provide higher returns than the other three stocks, but it can help explain that its returns are due to smarter, less volatile investments. Additionally, a negative Sharpe ratio, as observed by Vale S.A., can either mean one of two things: the return will be negative, or the chosen risk-free rate is greater than the return of the stock itself. Either way, its a good indication to invest your money elsewhere.

\
 
Now we consider consider an equally weighted portfolio of all of these stocks combined. We simply add each of the four stocks together per day, then take the average to calculate the equally weighted new stock. Below are the four stocks listed again with the "Combined" stock at the bottom for comparison:

```{r combined, include=TRUE, eval=TRUE, message=FALSE, warning=FALSE}

stocks2 <- as.xts(data.frame(VALE=dailyReturn(VALE),
                            CSCO=dailyReturn(CSCO),
                            INTC=dailyReturn(INTC),
                            WFC=dailyReturn(WFC)
                            ))

#creating a combined row with the averages for all companies
stocks2$Totals <- rowSums(stocks2)/4

stocks_combined <- cbind(data.frame(sapply(stocks2, mean)),
                         data.frame(sapply(stocks2, var))) 

colnames(stocks_combined) <- c("Mean", "Variance")
rownames(stocks_combined) <- c("Vale S.A.", "Cisco", "Intel", "Wells Fargo", "Combined")

```

```{r knitr3, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

if (!require("knitr")) {
     install.packages("knitr")
     library(knitr)
}

kable(stocks_combined)
```

Since they were equally weighted, the mean return of the Combined stock is just an average of the averages. So it would have a middle-of-the-road return, but with the lowest variance, it would also have the most predictable and least volatile return on average.

\

</div>

### References