---
title: "HW1_exer2"
author: "Group9"
date: "9/4/2020"
output: html_document
---

# Exercise 2: Financial Analysis

```{r trueinstall, echo=FALSE, message=FALSE, warning=FALSE}

if (!require("quantmod")) {
    install.packages("quantmod")
    library(quantmod)
}

```

The data set from four companies was obtained from yahoo, from January 1-2019 to September 6-2020. This was done utilizing the package `quantmod`:

```{r install, results=FALSE, message=FALSE, warning=FALSE}

start <- as.Date("2019-01-01")
end <- as.Date("2020-09-06")

getSymbols("VALE", src="yahoo", from=start, to=end)
getSymbols("CSCO", src="yahoo", from=start, to=end)
getSymbols("INTC", src="yahoo", from=start, to=end)
getSymbols("WFC", src="yahoo", from=start, to=end)

```

After that, all the data sets should be fetched and can be seen in the environment tab. We can take a look at them here using the command `head(dataset)`. For the VALE dataset, for example:

```{r viewdata, include=TRUE, eval=TRUE, echo=FALSE}

head(VALE)

```

An xts object `stocks` was created to unify all closing values from all the datasets, and then used to plot these values:

```{r plot, include=TRUE, eval=TRUE}

stocks <- as.xts(data.frame(VALE=VALE[, "VALE.Close"],
                            CSCO=CSCO[, "CSCO.Close"],
                            INTC=INTC[, "INTC.Close"],
                            WFC=WFC[, "WFC.Close"]
                            ))
plot(as.zoo(stocks),
     screens=1,
     lty=1:4,
     xlab="Date",
     ylab="Price",
     font.axis=1)
legend("topleft",
       inset=.02,
       c("VALE", "CSCO", "INTC", "WFC"),
       lty=1:4,
       cex=.5)
```

```{r candlechart, include=TRUE, eval=TRUE}

stocks2 <- as.xts(data.frame(VALE=VALE['2020-05-01/2020-09-06'],
                            CSCO=CSCO['2020-05-01/2020-09-06'],
                            INTC=INTC['2020-05-01/2020-09-06'],
                            WFC=WFC['2020-05-01/2020-09-06']
                            ))
candleChart(stocks2,
            up.col="gray",
            dn.col="blue",
            theme="white")


```



### 4
### 5
### 6
### 7 
